Exploring dimensions of biodiversity in Japanese ferns

### 2020-07-07 ms/SI.Rmd
Update MS

@@

**`r s_figure("grain-size")`**. `r s_figure_cap("grain-size")` **A**, Percent of grid cells with redundancy > 0.5 by grain size (0.1, 0.2, 0.3, or 0.4 degree grid-cells). **B**, Percent of grid cells with redundancy = 0 by grain size. **C**, Distribution of redundancy values by grain size.

@@

```{r k-plot, fig.width = 7, fig.height = 3}
k_taxonomy_select <- 8

a <- k_taxonomy$df %>%
  as_tibble() %>%
  mutate(
    optimal_auto = ifelse(k == k_taxonomy$optimal$k, TRUE, FALSE),
    optimal_manual = ifelse(k == k_taxonomy_select, TRUE, FALSE),
    label = ifelse(k == k_taxonomy_select, glue::glue("k = {k}"), "")
  ) %>%
  ggplot(aes(x = k, y = ev)) +
  geom_path() +
  geom_point(aes(color = optimal_manual, shape = optimal_auto), size = 2.5) +
  geom_text(aes(label = label), vjust = 0, nudge_y = 0.02, size = 4) +
  scale_color_manual(values = c("TRUE" = "#E31A1C", "FALSE" = "black")) +
  scale_shape_manual(values = c("TRUE" = 17, "FALSE" = 16)) +
  labs(
    y = "Explained variance",
    x = "Number of clusters") +
  theme(
    legend.position = "none"
  )

k_phylogeny_select <- 6

b <- k_phylogeny$df %>%
  as_tibble() %>%
  mutate(
    optimal_auto = ifelse(k == k_phylogeny$optimal$k, TRUE, FALSE),
    optimal_manual = ifelse(k == k_phylogeny_select, TRUE, FALSE),
    label = ifelse(k == k_phylogeny_select, glue::glue("k = {k}"), "")
  ) %>%
  ggplot(aes(x = k, y = ev)) +
  geom_path() +
  geom_point(aes(color = optimal_manual, shape = optimal_auto), size = 2.5) +
  geom_text(aes(label = label), vjust = 0, nudge_y = 0.02, size = 4) +
  scale_color_manual(values = c("TRUE" = "#E31A1C", "FALSE" = "black")) +
  scale_shape_manual(values = c("TRUE" = 17, "FALSE" = 16)) +
  labs(
    y = "Explained variance",
    x = "Number of clusters",
    shape = "Optimal k\n(auto)",
    color = "Optimal k\n(manual)") +
  theme(
    axis.title.y = element_blank()
  )

@@

**`r s_figure("k-plot")`**. `r s_figure_cap("k-plot")` **A**, taxonomic bioregions. **B**, phylogenetic bioregions. Shape indicates optimal *k* selected by \function{optimal{\_}phyloregion} function in the R package \package{phyloregion} (circle, non-optimal; triangle, optimal); color indicates optimal *k* selected manually (black, non-optimal; red, optimal; see Methods).

>>>>

```{r sampling-curve, fig.width = 4, fig.height = 3}
# Format fern abundances (no hybrids) as numeric vector
fern_abun <-
  occ_point_data_ferns %>%
  group_by(taxon) %>%
  count(sort = TRUE) %>%
  pull(n)

# Run iNEXT
inext_out <- iNEXT::iNEXT(fern_abun, q=0, datatype="abundance")

# create subset of observed points to add to plot as dots
observed_data <- inext_out$iNextEst[inext_out$iNextEst$method == "observed",]

# check estimated completeness
est_completeness <- inext_out$iNextEst %>% filter(method == "observed") %>% pull(SC) %>% scales::percent()
rich_interval <- glue::glue("{observed_data$qD.LCL %>% round} to {observed_data$qD.UCL %>% round}")
completeness_interval <- glue::glue("{observed_data$SC.LCL %>% scales::percent()} to {observed_data$SC.UCL %>% scales::percent()}")

# Make iNEXT plot
inext_out$iNextEst %>%
  filter(method != "observed") %>%
  ggplot(aes(x=m, y=qD)) +
  geom_ribbon(aes(ymin = qD.LCL, ymax = qD.UCL), fill = "grey70") +
  geom_line(aes(linetype=method), size=0.8) +
  scale_linetype_manual(values=c("dotted", "solid", "solid")) +
  geom_point(data=observed_data, size=3) +
  scale_y_continuous("Richness") +
  scale_x_continuous("No. specimens") +
  theme_minimal() +
  theme(
    legend.position="bottom", 
    legend.title=element_blank()
  )

ggsave(result_file(s_figure("sampling-curve"), "pdf"), width = 4, height = 3, units = "in")
```

**`r s_figure("sampling-curve")`**. `r s_figure_cap("sampling-curve")` Grey shading indicates 95% confidence interval (not visible where it does not extend beyond the fitted line). Point indicates observed taxonomic richness (`r observed_data$qD` taxa, 95% confidence interval `r rich_interval` taxa). Taxa include species, subspecies, and varieties, excluding hybrids. Sampling completeness is estimated to be `r est_completeness` (95% confidence interval `r completeness_interval`).

@@

**`r s_figure("grain-size")`**. `r s_figure_cap("grain-size")` **A**, Percent of grid cells with redundancy > 0.5 by grain size (0.1, 0.2, 0.3, or 0.4 degree grid-cells). **B**, Percent of grid cells with redundancy = 0 by grain size. **C**, Distribution of redundancy values by grain size. There is a sudden improvement in redundancy values from 0.1 to 0.2 degree grid-cells, but less improvement as grain size is increased beyond 0.2 degrees.

@@

```{r k-plot, fig.width = 7, fig.height = 3}
a <- k_taxonomy$df %>%
  as_tibble() %>%
  mutate(
    optimal_auto = ifelse(k == k_taxonomy$optimal$k, TRUE, FALSE),
    label = ifelse(k == k_taxonomy$optimal$k, glue::glue("k = {k}"), "")
  ) %>%
  ggplot(aes(x = k, y = ev)) +
  geom_path() +
  geom_point(aes(color = optimal_auto), size = 2.5) +
  geom_text(aes(label = label), vjust = 0, nudge_y = 0.02, size = 4) +
  scale_color_manual(values = c("TRUE" = "#E31A1C", "FALSE" = "black")) +
  labs(
    y = "Explained variance",
    x = "Number of clusters") +
  theme(
    legend.position = "none"
  )

k_phylogeny_select <- 8

b <- k_phylogeny$df %>%
  as_tibble() %>%
  mutate(
    optimal_auto = ifelse(k == k_phylogeny$optimal$k, TRUE, FALSE),
    label = ifelse(k == k_phylogeny$optimal$k, glue::glue("k = {k}"), "")
  ) %>%
  ggplot(aes(x = k, y = ev)) +
  geom_path() +
  geom_point(aes(color = optimal_auto), size = 2.5) +
  geom_text(aes(label = label), vjust = 0, nudge_y = 0.02, size = 4) +
  scale_color_manual(values = c("TRUE" = "#E31A1C", "FALSE" = "black")) +
  labs(
    y = "Explained variance",
    x = "Number of clusters") +
  theme(
    legend.position = "none"
  )

@@

**`r s_figure("k-plot")`**. `r s_figure_cap("k-plot")` **A**, taxonomic bioregions. **B**, phylogenetic bioregions. Color indicates optimal *k* selected selected by \function{optimal{\_}phyloregion} function in the R package \package{phyloregion} (black, non-optimal; red, optimal; see Methods).

### 2020-07-07 ms/manuscript.Rmd
Update MS

loadd(list = c(
  "occ_point_data_ferns",
  "ses_phy_ferns",
  "ses_phy_ferns_endemic",
  "ses_traits_ferns",
  "shape_ferns",
  "k_taxonomy",
  "k_phylogeny",
  "regions_taxonomy",
  "regions_phylogeny"
),
cache = ja_fern_cache)

@@

Here we leverage this exceptionally rich dataset to analyze the geographic distribution of biodiversity, reveal its spatial structure, and assess drivers of community assembly.

@@

We used occurrence data of ferns in Japan comprising `r nrow(occ_point_data_ferns) %>% scales::number(big.mark = ",")` vouchered specimens representing `r occ_point_data_ferns %>% pull(taxon) %>% n_distinct()` non-hybrid taxa (species, subspecies, and varieties). All specimens are housed at herbaria in Japan, and taxonomy follows the Japan Green List v. 1.01 (http://www.rdplants.org/gl/).

Any analysis of biodiversity must consider that smaller grain size is needed to detect environmental effects, while larger grain size is needed to ensure adequate species sampling. We tested using 0.1, 0.2, or 0.4 degree grid-cells for defining communities (co-occurring species). At each size, species occurrences were converted into a presence-absence community matrix (a species was considered present if at least one specimen was recorded in that grid cell). We calculated sampling redundance (1 - richness/number of specimens) [@Garcillan2003] to quantify adequacy of sampling. Preliminary analysis indicated that 0.2 degree grid-cells are optimal for our dataset, so we used this for all subsequent analyses.

@@

We used ploidy level and breeding system data in @Ebihara2019b.

@@

We measured taxanomic diversity using species richness.

@@

### Analysis of bioregions and ecological structure

>>>>

loadd(list = c(
  "occ_point_data_ferns",
  "occ_point_data_ferns_unfiltered",
  "ses_phy_ferns",
  "ses_phy_ferns_endemic",
  "ses_traits_ferns",
  "shape_ferns",
  "k_taxonomy",
  "k_phylogeny",
  "regions_taxonomy",
  "regions_phylogeny"
),
cache = ja_fern_cache)

@@

Here we leverage this exceptionally rich dataset to analyze the geographic distribution of biodiversity in detail. We ask the following questions in our study system, the ferns of Japan: 1. How is biodiversity distributed? 2. How is biodiversity structured? 3. How well is biodiversity protected?

@@

We used a list of all non-hybrid fern specimens housed at herbaria in Japan. We chose to use this over other sources (e.g., GBIF), because of its high taxonomic quality: all identifications are standardized to a common taxonomy, the Japan Green List v. 1.01 (http://www.rdplants.org/gl/). Furthermore, the overall sampling completeness of this dataset is extremely high (`r s_figure("sampling-curve")`). The original list comprises `r nrow(occ_point_data_ferns_unfiltered) %>% scales::number(big.mark = ",")` specimens representing `r occ_point_data_ferns_unfiltered %>% pull(taxon) %>% n_distinct()` taxa (species, subspecies, and varieties). We filtered out any occurrences not within a map of standard grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html). The final cleaned occurrence dataset included `r nrow(occ_point_data_ferns) %>% scales::number(big.mark = ",")` specimens (`r occ_point_data_ferns %>% pull(taxon) %>% n_distinct()` taxa).

A necessary step in any analysis of biodiversity is to set the grain size used to define communities. This decision must consider that smaller grain size is needed to detect environmental effects, while larger grain size is needed to ensure adequate species sampling. We tested using 0.1, 0.2, 0.3, and 0.4 degree grid-cells for defining communities (co-occurring species). At each size, species occurrences were converted into a presence-absence community matrix (a species was considered present if at least one specimen was recorded in that grid cell). We calculated sampling redundance (1 - richness/number of specimens) [@Garcillan2003] to quantify adequacy of sampling. Preliminary analysis indicated that 0.2 degree grid-cells are optimal for our dataset (`r s_figure("grain-size")`). We filtered out any cells with redundancy < 0.1 (mostly cells along the coast with very low land area) and used this for subsequent analyses.

@@

@@

We measured taxonomic diversity using taxonomic richness.

@@

### Analysis of bioregions
